Asymptotic and Numerical Solution for Electro-Kinetic Flow Around a Spherical Particle

### 2012-09-11 notes/work/ijnam.tex
work on LaTeX

\subsubsection  {Bulk-scale equations}
The Nernst-Planck equations above \eqref{eq:fluxes} for the fluid bulk (outside
the boundary layer), using bulk variables denoted by uppercase letters, read:
\begin{eqnarray} \begin{array}{cccc} \label{eq:bulk_flux}
  Q = 0, &
  C = C_+ = C_-, &
\bJ = -\bnabla C + \alpha \bV C, &
\bI = -C \bnabla \varPhi.
\end{array}\end{eqnarray}
Because the flow is incompressible ($\bnabla \cdot V = 0$), 
ion conservation results in the following 
salt and charge conservation equations -- the salt and charge fluxes above 
\eqref{eq:bulk_flux} are divergence-free
\eqref{eq:zero_flux}:
\begin{eqnarray} 
\label{eq:salt_charge}
\Laplacian C - \alpha \bV \cdot \bnabla C = 0, &
\bnabla \cdot \pars{ C \bnabla \varPhi } = 0.
\end{eqnarray}
The macroscale Stokes flow equations have the same form as before \eqref{eq:stokes}:
\begin{eqnarray} 
\bnabla \cdot \bV = 0, &   
\Laplacian \bV - \bnabla P + \bnabla \varPhi \Laplacian \varPhi = \bzero.
\end{eqnarray}

@@

\subsection{System Equations} \label{sec:disc_equations}
Charge and salt fluxes \eqref{eq:bulk_flux} are discretized on grid cell boundaries 
(using upwind scheme $\mathcal{U}$ for numerical stability at large cell Peclet number):
\begin{eqnarray} \nonumber
I^h_r = -\cI_r(C^h) \cdot \cD_r(\varPhi^h), &
I^h_\theta = -\cI_\theta(C^h) \cdot \frac{\cD_\theta(\varPhi^h)}{r}, \\
J^h_r = -\cD_r(C^h) + \alpha V^h_r \cdot \mathcal{U}^{\bV^h}_r (C^h), &
J^h_\theta = -\frac{\cD_\theta(C^h)}{r} + \alpha V^h_\theta \cdot \mathcal{U}^{\bV^h}_\theta (C^h), 
\\ \nonumber
 \mathcal{U}^{\bV^h}_r(C^h)[i,j] = C^h\left[i-\frac{\sign(V^h_r)}{2}, j\right], &
 \mathcal{U}^{\bV^h}_\theta(C^h)[i,j] = C^h\left[i, j-\frac{\sign(V^h_\theta)}{2}\right]. 
\end{eqnarray}

@@

Force components ($\bF = -\bnabla P + \bLaplacian \bV + \bnabla \varPhi \Laplacian \varPhi$) 
are discretized on the staggered velocity grid, where
linear interpolation is used for the Coulomb force and for vector Laplacian components.
\begin{eqnarray}
\nonumber
F^h_r &=& -\cD_r(P^h) 
          + \cL(V^h_r) - \frac{2}{r^2} V^h_r 
		  - \frac{2}{r^2 \sin\theta} \cI_r(\cD_\theta (V^h_\theta \sin\theta))
          + \cD_r(\varPhi^h) \cdot \cI_r(\cL(\varPhi^h)), \\
F^h_\theta &=& -\frac{\cD_\theta(P^h)}{r} 
		  + \cL(V^h_\theta) - \frac{F^h_\theta}{r^2 \sin^2\theta} 
		  + \frac{2}{r^2} \cI_\theta(\cD_\theta(F^h_r))
		  + \frac{\cD_\theta(\varPhi^h)}{r} \cdot \cI_\theta(\cL(\varPhi^h)), \\
\nonumber
\cL(f^h) &=& \frac{1}{r^2}\cD_r\pars{\cD_r(f^h) r^2} + 
\frac{1}{r^2 \sin\theta} \cD_\theta\pars{\cD_\theta(f^h) \cdot \sin\theta}.
\end{eqnarray}

@@

\subsubsection{Operator Representation}
System variable $\bx$ is defined as the concatenation $\bx = [\,\varPhi, C, V_r, V_\theta, P\,]$.
Thus, each problem variable can be computed by applying an appropriate projection operator on $\bx$.
The system can be written as a nonlinear operator $\cO(\bx) = \bzero$ 
acting on the system variable ${\bx}$, 
where $\cO$ is the system equations operator, 
derived by concatenation of the system equations (described at Subsection \ref{sec:disc_equations})
and the boundary conditions operators (described at Subsection \ref{sec:disc_boundary}).
This way, the same numerical solver handles both the nonlinear system equation and nonlinear
boundary conditions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Asymptotic Analysis} \label{sec:asymp}
In order to validate the numerical results, an asymptotic analysis is applied 
to the nonlinear systems of ion-exchanger migration (\cite{yariv2010migration,zeyde2012report})
and highly charged particle electrophoresis (\cite{schnitzer2012surface, schnitzer2012cubic})

The nonlinear system is written as $\cO(\bx;\beta) = \bzero$.
For small $\beta$, the solution to the nonlinear system can be expanded in a Taylor series in $\beta$:
\begin{eqnarray*}
\bx = \bx(\beta) &\approx& \sum_n \bx_n \beta^n.
\end{eqnarray*}
Thus, the system variables are expanded around $\beta = 0$
(where the velocity $\bV$ and pressure $P$ are computed from the streamfunction $\Psi$):
\begin{eqnarray*}
\varPhi = \sum_i \beta^i \varPhi_i, &
C = 1 + \sum_i \beta^i C_i, &
\Psi = \sum_i \beta^i \Psi_i,
\end{eqnarray*}

>>>>

\subsubsection  {Bulk-scale equations}
The Nernst-Planck equations above \eqref{eq:fluxes} for the fluid bulk (outside
the boundary layer), using bulk variables denoted by uppercase letters, read:
\begin{eqnarray*}
\begin{array}{cccc} \label{eq:bulk_flux}
  Q = 0, &
  C = C_+ = C_-, &
\bJ = -\bnabla C + \alpha \bV C, &
\bI = -C \bnabla \varPhi.
\end{array}\end{eqnarray*}
Because the flow is incompressible ($\bnabla \cdot V = 0$), 
ion conservation results in the following 
salt and charge conservation equations -- the salt and charge fluxes above 
\eqref{eq:bulk_flux} are divergence-free
\eqref{eq:zero_flux}:
\begin{eqnarray} 
\label{eq:salt_charge}
\Laplacian C - \alpha \bV \cdot \bnabla C = 0, &
\bnabla \cdot \pars{ C \bnabla \varPhi } = 0.
\end{eqnarray}
The macroscale Stokes flow equations have the same form as before \eqref{eq:stokes}:
\begin{eqnarray} 
\bnabla \cdot \bV = 0, &   
\Laplacian \bV - \bnabla P + \bnabla \varPhi \Laplacian \varPhi = \bzero.
\end{eqnarray}

@@

\subsection{System Equations} \label{sec:disc_equations}

Ion flux ($\bI = - C \bnabla \varPhi$) is discretized on grid cell boundaries:
\begin{eqnarray*}
I^h_r = -\cI_r(C^h) \cdot \cD_r(\varPhi^h), &
I^h_\theta = -\cI_\theta(C^h) \cdot \frac{\cD_\theta(\varPhi^h)}{r}.
\end{eqnarray*}

Salt flux ($\bJ = -\bnabla C + \alpha \bV C$) is discretized on grid cell boundaries 
(using upwind scheme $\mathcal{U}$ for numerical stability at large cell Peclet number):
\begin{eqnarray*}
J^h_r = -\cD_r(C^h) + \alpha V^h_r \cdot \mathcal{U}^{\bV^h}_r (C^h), &
J^h_\theta = -\frac{\cD_\theta(C^h)}{r} + \alpha V^h_\theta \cdot \mathcal{U}^{\bV^h}_\theta (C^h), 
\\
 \mathcal{U}^{\bV^h}_r(C^h)[i,j] = C^h\left[i-\frac{\sign(V^h_r)}{2}, j\right], &
 \mathcal{U}^{\bV^h}_\theta(C^h)[i,j] = C^h\left[i, j-\frac{\sign(V^h_\theta)}{2}\right]. 
\end{eqnarray*}

@@

Force components ($\bF = -\bnabla P + \bLaplacian \bV + \bnabla \varPhi \Laplacian \varPhi$) 
are discretized on the staggered velocity grid, where
linear interpolation is used for the Coulomb force and for vector Laplacian components.
\begin{eqnarray*}
F^h_r &=& -\cD_r(P^h) 
          + \cL(V^h_r) - \frac{2}{r^2} V^h_r 
		  - \frac{2}{r^2 \sin\theta} \cI_r(\cD_\theta (V^h_\theta \sin\theta))
          + \cD_r(\varPhi^h) \cdot \cI_r(\cL(\varPhi^h)), \\
F^h_\theta &=& -\frac{\cD_\theta(P^h)}{r} 
		  + \cL(V^h_\theta) - \frac{F^h_\theta}{r^2 \sin^2\theta} 
		  + \frac{2}{r^2} \cI_\theta(\cD_\theta(F^h_r))
		  + \frac{\cD_\theta(\varPhi^h)}{r} \cdot \cI_\theta(\cL(\varPhi^h)), \\
\cL(f^h) &=& \frac{1}{r^2}\cD_r\pars{\cD_r(f^h) r^2} + 
\frac{1}{r^2 \sin\theta} \cD_\theta\pars{\cD_\theta(f^h) \cdot \sin\theta}.
\end{eqnarray*}

@@

\subsubsection{Operator Representation}
System variable $\bx$ is defined as the concatenation $\bx = [\,\varPhi, C, V_r, V_\theta, P\,]$.
Thus, each problem variable can be computed by applying an appropriate projection operator on $\bx$:
\begin{eqnarray*}
X &=& \cP_X(\bx) \text{ for } X \in \left\{\varPhi, C, V_r, V_\theta, P\right\}.
\end{eqnarray*}
The system can be written as a nonlinear operator $\cO(\bx) = \bzero$ 
acting on the system variable ${\bx}$, 
where $\cO$ is the system equations operator, 
derived by concatenation of the system equations (described at Subsection \ref{sec:disc_equations})
and the boundary conditions operators (described at Subsection \ref{sec:disc_boundary}).
This way, the same numerical solver handles both the nonlinear system equation and nonlinear
boundary conditions.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Asymptotic Analysis} \label{sec:asymp}
The nonlinear system is defined by the following PDEs:
\label{PDEs}
\begin{eqnarray*}
\begin{array}{cccc}
\bnabla \cdot \pars{C \bnabla \varPhi} = 0, &
\bnabla \cdot \pars{\bnabla C - \alpha \bV C} = 0, &
\bnabla \cdot \bV = 0, &
\bLaplacian \bV - \bnabla P + \Laplacian \varPhi \bnabla \varPhi = \bzero.
\end{array}\end{eqnarray*}

For $\zeta(\theta) = \cV - \varPhi(1, \theta)$ and 
$\cV = -\log \gamma$, the boundary conditions at $r = 1$ are given by
\begin{eqnarray*}
\begin{array}{rcl}
\varPhi + \log C &=& \cV + \log \gamma = 0, \\
\deriv{}{r}\pars{\varPhi - \log C} &=& \frac{J_-}{C} = 0, \\
\bV = \zeta \cdot \bnabla_\mathcal{S} \varPhi 
+ 2\log\pars{1-\tanh^2\frac{\zeta}{4}} \cdot \bnabla_\mathcal{S} \log C
&=& 4\log\pars{\frac{1 + e^\frac{\zeta}{2}}{2}} \bnabla_S \varPhi.
\end{array}
\end{eqnarray*}

The boundary conditions for $r \rightarrow \infty$ are given by
\begin{eqnarray*}
\bnabla \varPhi = -\beta \ui, &
C = 1, &
\bV = -\cU \ui,
\end{eqnarray*}
and the total force acting on the particle is zero: $\bF = \bzero$.

The nonlinear system can be written as: 
\begin{eqnarray*}
\cO(\bx;\beta) &=& \bzero.
\end{eqnarray*}

For small $\beta$, the solution to the nonlinear system can be expanded in a Taylor series in $\beta$.
\begin{eqnarray*}
\bx = \bx(\beta) &\approx& \sum_n \bx_n \beta^n.
\end{eqnarray*}

All the variables are expanded around $\beta = 0$:
\begin{eqnarray*}
\varPhi = \sum_i \beta^i \varPhi_i, &
C = 1 + \sum_i \beta^i C_i, &
\Psi = \sum_i \beta^i \Psi_i,
\end{eqnarray*}
where the velocity $\bV$ and pressure $P$ are computed from the streamfunction $\Psi$:
\begin{eqnarray*}
\bV &=& \bnabla \times \pars{\frac{\Psi}{r \sin\theta} \bphihat} 
= \frac{1}{r^2 \sin\theta} \deriv{\Psi}{\theta} \brhat 
- \frac{1}{r \sin\theta} \deriv{\Psi}{r} \bthetahat.
\end{eqnarray*}
